---
title: "weight analysis"
author: "Steve Rolfe"
date: "27 October 2015"
output: html_document
---

Standard code to clear the system and install required libraries
```{r}
rm(list=ls())
library(ggplot2)
library(lubridate)
library(dplyr)
library(zoo)

#load the multiplot function from the R cookbook from the R local directory
pathnames <- list.files(pattern="[.]R$", path="./R/", full.names=TRUE)
pathnames
sapply(pathnames, FUN=source)

```


```{r}
#plotting parameters for saving plots - output from ggplot
#ggsave(file="filename.pdf",path="./output")
#after calling ggplot

#output from multiplot
#pdf(file="filename.pdf")
#before calling multiplot
#dev.off()


```

Set up standard files names and the subset to plot
```{r}
setwd("C:/Users/Steve Rolfe/Google Drive/Phenomics/Weight_analysis")

input_file="./data/BR7_final_weights.csv"
output_file="./output/BR7_weight_out.csv"
plant_codes="./data/BR7_plant_codes.csv"

#subset of data to plot using grep format
sset<-"BR7-0[1-5]1[1-3][1-4]"
```

Set up standard headers for the files - these need to be edited to be consistent
The BR codes are suppled by IBERS but we should include the ASSYT codes where possible
Standard headers should be
BR_code,date_time, measurements (with informative names)
Custom code will be needed to translate the BR_codes into useful information

Read in the data and check that the required headers are there
```{r}
#read in the data
wtdata<-read.csv(input_file,header=TRUE,stringsAsFactors = FALSE)
if(is.object(wtdata)==FALSE) {stop(c("Can't open file:",input_file))}

#check that there the required headings are there and throw an error if not
if(is.na(match("BR_code",colnames(wtdata)))){stop("BR_code missing from input file header")}
if(is.na(match("date_time",colnames(wtdata)))){stop("date_time missing from input file header")}
head(wtdata)

#this file contains info about the plants used
plant_codes<-read.csv(plant_codes,header = TRUE,stringsAsFactors = FALSE)
if(is.object(plant_codes)==FALSE) {stop(c("Can't open file:",plant_codes))}
```

We  deal with any specific coding here about names of plants, spores etc.
Use the file plant_codes to keep the information organised
```{r}
#dpi offset from first data point
#it's good to express everything as dpi but 1st data point might not be dpi
#set the date of infection
dpi_datetime<-"2015-7-6 00:00:00"

#now set the factors manually
wtdata$BR_code<-as.factor(wtdata$BR_code)

#and set the plants,spores,etc
#the plant code is the sixth character in the BR7 string
wtdata$plant<-as.factor(substr(wtdata$BR_code,6,6))
#the spore concentration code is the 8th character
wtdata$spore<-as.factor(substr(wtdata$BR_code,8,8))
#the repeat is the 9th
wtdata$rep<-as.factor(substr(wtdata$BR_code,9,9))
head(wtdata)

#now merge the plant names with the data frame
wtdata<-merge(wtdata,plant_codes,by.x="plant",by.y="plant_code")
head(wtdata)

#now do some standard time processing
#use difftime to get the days since inoculation
wtdata$dpi<-difftime(wtdata$date_time,dpi_datetime,units="days")
#make it a factor when we just have a few days (but not when we want a continuous x scale)
wtdata$dpi<-as.factor(wtdata$dpi)
```

Plot the data 
Always use a subset function to keep things under control for plotting
```{R}
subdata<-subset(wtdata,grepl(sset,wtdata$BR_code))
subdata<-droplevels(subdata)
#plot the raw data as a boxplot
p1<-ggplot(subdata,aes(x=spore,ShootFW))
p1<-p1+geom_boxplot(aes(fill=spore),outlier.colour="red",outlier.size=3)
p1<-p1+facet_wrap(~plant_name,nrow=1)
p1

p2<-ggplot(subdata,aes(x=spore,RootFW))
p2<-p2+geom_boxplot(aes(fill=spore),outlier.colour="red",outlier.size=3)
p2<-p2+facet_wrap(~plant_name,nrow=1)
p2


#pdf(file="./output/filename.pdf")
multiplot(p1,p2)
#dev.off()

p3<-ggplot(subdata,aes(x=spore,ShootDW))
p3<-p3+geom_boxplot(aes(fill=spore),outlier.colour="red",outlier.size=3)
p3<-p3+facet_wrap(~plant_name,nrow=1)
p3

p4<-ggplot(subdata,aes(x=spore,RootDW))
p4<-p4+geom_boxplot(aes(fill=spore),outlier.colour="red",outlier.size=3)
p4<-p4+facet_wrap(~plant_name,nrow=1)
p4

multiplot(p3,p4)
```
Calculate root:shoot ratios
```{R}
subdata$RSDW<-subdata$RootDW/subdata$ShootDW
subdata$RSFW<-subdata$RootFW/subdata$ShootFW
subdata$biomass_allocDW<-subdata$RootDW/(subdata$RootDW+subdata$ShootDW)
subdata$biomass_allocFW<-subdata$RootFW/(subdata$RootFW+subdata$ShootFW)

p5<-ggplot(subdata,aes(x=spore,RSDW))
p5<-p5+geom_boxplot(aes(fill=spore),outlier.colour="red",outlier.size=3)
p5<-p5+facet_wrap(~plant_name,nrow=1)
p5

p6<-ggplot(subdata,aes(x=spore,biomass_allocDW))
p6<-p6+geom_boxplot(aes(fill=spore),outlier.colour="red",outlier.size=3)
p6<-p6+facet_wrap(~plant_name,nrow=1)
p6
multiplot(p5,p6)

p7<-ggplot(subdata,aes(x=spore,RSFW))
p7<-p7+geom_boxplot(aes(fill=spore),outlier.colour="red",outlier.size=3)
p7<-p7+facet_wrap(~plant_name,nrow=1)
p7

p8<-ggplot(subdata,aes(x=spore,biomass_allocFW))
p8<-p8+geom_boxplot(aes(fill=spore),outlier.colour="red",outlier.size=3)
p8<-p8+facet_wrap(~plant_name,nrow=1)
p8
multiplot(p7,p8)

write.csv(subdata,file=output_file)
```